The value of being indispensable: a cooperative approach to bargaining with a continuum of players

### 2024-05-23 src/paper/application.tex
Spelling

Nevertheless, there is an important distinction compared to the previous results.
While this does lead to a decrease in total surplus, in contrast to the two-sided bargaining case, this does not imply a decrese in consumer surplus.
As mentioned before, the consumers' share of total surplus is equal to the platform's profits, and thus it is increasing in $N_P$ if and only if platform profits (\cref{fig:profits_full_surplus_two_sided}) are.
Therefore, consumers might prefer hybrid operation to the platform being a pure marketplace.

@@

In words, \cref{prop:outcome_based_bargaining_power} describes two alternative worlds, in which $N_F$ fringe firms (assuming a platform with $N_P$ product variety and $N_C$ consumers) are able to achieve the same total profit level.
However, in one of these cases, the fringe firms are more substitutable to each other in the sense that fewer of them are needed to achieve a given level of profit (see \cref{fig:outcome_based_bargaining_power}).
The observation is that, in this situation, the platform's share is indeed higher when the fringe firms are more substitutable.
This coincides with the intuitive idea that the platform's bargaining power is higher when it does not really mind loosing a few fringe sellers.

@@

Consider the game $\mathcal{G} = (\mathcal{N}, v)$, characterized by the set of players $\mathcal{N}$ and its characteristic function $v$.
The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i\}, i \in [0, N_F]$.
Because the platform is indispensible, the value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
\begin{align*}
    v(S) = \begin{cases}
        0 & \text{if } P \notin S \\
        N_C f(N_P, n_F(S)) & \text{if } P \in S
    \end{cases},
\end{align*}
where $n_F(S)$ is the measure of fringe firms in $S$.

@@

        \node[align=center, right=of entry_decision, distance=3cm] (arrange_fringe) {Arrange fringe firms in some fixed order};
        \node[align=center, below of=arrange_fringe] (foreach) {For each (remaining) fringe firm $F_i$:};
        \node[align=center] (choose_proposer) [below=1cm of foreach] {$F_i$ or $P$ is choosen as the proposer};
        \node[align=center] (proposal) [below of=choose_proposer] {Proposer proposes entry fee $K_{F_i}$};
        \node[align=center] (accept) [below left=2cm and -2cm of proposal] {Responder accepts $K_{F_i}$};
        \node[align=center] (reject) [below right=2cm and -2cm of proposal] {Responder rejects $K_{F_i}$};
        \node[align=center] (next_player) [below of=accept] {Move on to the \\ next fringe firm};
        \node[align=center] (eliminate) [below of=reject] {Eliminate $F_i$};

@@

\section{The platform's product pricing assumption}
\label{sec:separate_pricing}

TODO: discuss the assumptioin of pricing products separately

>>>>

Nevertheless, there is an important distinction compared to the previous results.
While this does lead to a decrease in total surplus, in contrast to the two-sided bargaining case, this does not imply a decrease in consumer surplus.
As mentioned before, the consumers' share of total surplus is equal to the platform's profits, and thus it is increasing in $N_P$ if and only if platform profits (\cref{fig:profits_full_surplus_two_sided}) are.
Therefore, consumers might prefer hybrid operation to the platform being a pure marketplace.

@@

In words, \cref{prop:outcome_based_bargaining_power} describes two alternative worlds, in which $N_F$ fringe firms (assuming a platform with $N_P$ product variety and $N_C$ consumers) are able to achieve the same total profit level.
However, in one of these cases, the fringe firms are more substitutable to each other in the sense that fewer of them are needed to achieve a given level of profit (see \cref{fig:outcome_based_bargaining_power}).
The observation is that, in this situation, the platform's share is indeed higher when the fringe firms are more substitutable.
This coincides with the intuitive idea that the platform's bargaining power is higher when it does not really mind losing a few fringe sellers.

@@

Consider the game $\mathcal{G} = (\mathcal{N}, v)$, characterized by the set of players $\mathcal{N}$ and its characteristic function $v$.
The set of players consists of the platform plus the fringe firms: $\mathcal{N} = \{P, F_i\}, i \in [0, N_F]$.
Because the platform is indispensable, the value of a coalition $S$ is zero without the platform, and depends on the number of fringe firms otherwise:
\begin{align*}
    v(S) = \begin{cases}
        0 & \text{if } P \notin S \\
        N_C f(N_P, n_F(S)) & \text{if } P \in S
    \end{cases},
\end{align*}
where $n_F(S)$ is the measure of fringe firms in $S$.

@@

        \node[align=center, right=of entry_decision, distance=3cm] (arrange_fringe) {Arrange fringe firms in some fixed order};
        \node[align=center, below of=arrange_fringe] (foreach) {For each (remaining) fringe firm $F_i$:};
        \node[align=center] (choose_proposer) [below=1cm of foreach] {$F_i$ or $P$ is chosen as the proposer};
        \node[align=center] (proposal) [below of=choose_proposer] {Proposer proposes entry fee $K_{F_i}$};
        \node[align=center] (accept) [below left=2cm and -2cm of proposal] {Responder accepts $K_{F_i}$};
        \node[align=center] (reject) [below right=2cm and -2cm of proposal] {Responder rejects $K_{F_i}$};
        \node[align=center] (next_player) [below of=accept] {Move on to the \\ next fringe firm};
        \node[align=center] (eliminate) [below of=reject] {Eliminate $F_i$};

@@

% \section{The platform's product pricing assumption}
% \label{sec:separate_pricing}

% This section explores the consequences of relaxing the assumption that the platform prices its own products as if they were priced by separate entities.
% So let us now assume, that when pricing product $P_i$, the platform does consider its effect on the sales of its other products.

% For the fringe firms, the problem is the same as before: they choose their prices to maximize their profits, while not having an effect on total aggregate.
% Thus, the price they set is the same as before: $p_{F_i} = c_{F_i} + \mu$.

% The same is not the case with the platform.
% When pricing its products, the platform must consider the effect of its prices on the sales of its other products, through its effects on the aggregate.
% The end result is that the platform will set higher prices than under the separate pricing assumption, because it internalizes the effect of one product's price on the demand for the other products.
% In particular, it can be shown that the optimal price for the platform's product is
% \begin{align*}
%     p_P^* = c_P + \mu + W_0\left( \frac{N_P V_P}{N_F V_F + 1} \right),
% \end{align*}
% where $W_0$ is the principal branch of the Lambert $W$ function.

% Given an aggregate of size $A$, the platform's profits are
% \begin{align*}
%     \dots
% \end{align*}

### 2024-05-23 src/paper/theory.tex
Spelling

For practical applications, the case of one indispensable player and a large number of small ones is particularly important in a number of settings.
Examples include intra-firm ownership strutures \parencite{hart1990property}, wage bargaining \parencite{stole1996intra,stole1996organizational,levy1997individual}, collusion and mergers \parencite{segal2003collusion} and upstream-downstream supply chains \parencite{inderst2003bargaining,montez2007downstream}.
The study of multi-sided markets (or platforms) is also a good fit for this modeling approach, as the platform is often assumed to be indispensable for the various sides to interact.\footnote{
    \textcite{huang2022shapley} is amongst the very few examples of using cooperative game theory in this setting.
}

>>>>

For practical applications, the case of one indispensable player and a large number of small ones is particularly important in a number of settings.
Examples include intra-firm ownership structures \parencite{hart1990property}, wage bargaining \parencite{stole1996intra,stole1996organizational,levy1997individual}, collusion and mergers \parencite{segal2003collusion} and upstream-downstream supply chains \parencite{inderst2003bargaining,montez2007downstream}.
The study of multi-sided markets (or platforms) is also a good fit for this modeling approach, as the platform is often assumed to be indispensable for the various sides to interact.\footnote{
    \textcite{huang2022shapley} is amongst the very few examples of using cooperative game theory in this setting.
}
